
\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[backend=biber, style=authortitle, citestyle=verbose-trad1]{biblatex}
\usepackage{biblatex}
\addbibresource{Philosophical_Essay_References.bib}

\title{Ethics Rooted in Reality: Exploring Naturalistic Realism in Metaethics}
\author{Daniel Schellhorn}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}
In the realm of philosophical inquiry, few topics have spurred as much introspection and debate as the nature of ethics and its foundational underpinnings. The exploration of metaethics, particularly through the lens of naturalistic realism, offers a fascinating convergence of moral philosophy with the empirical understanding of the world. This essay delves into the intricate framework of naturalistic realism within metaethics as presented in chapter 2 of the \citetitle{mcpherson2020metaethics} \footcite{mcpherson2020metaethics}, a standpoint that endeavors to reconcile ethical thought and practice with our broader comprehension of the natural universe.

Metaethics, by its very definition, transcends the simplicity of distinguishing right from wrong, delving instead into the fundamental nature of morality itself. It questions the origins of moral principles and investigates the essence of what constitutes ethical reasoning. Within this complex landscape, naturalistic realism emerges as a pivotal doctrine, asserting that ethical truths are grounded in the natural world, devoid of any supernatural or non-natural dimensions. This perspective challenges the traditional dichotomy between facts and values, suggesting a more integrated approach where ethical principles are not only derived from but are also coherent with the observable universe.

The significance of naturalistic realism in metaethics lies in its attempt to address some of the most profound and enduring questions in moral philosophy. These include the metaphysical quandary of how values fit into a world predominantly understood through factual knowledge, the conceptual relationship between normative claims (what ought to be) and descriptive claims (what is), the epistemological challenge of justifying ethical beliefs that often stem from intuitive judgments, and the semantic exploration of how our language and thoughts connect with ethical properties or facts. By proposing a unified response to these multifaceted challenges, naturalistic realism seeks to offer a comprehensive account of ethics that is deeply rooted in the natural world.

The philosophical journey through naturalistic realism is not merely an academic exercise; it holds profound implications for our understanding of ethics in the contemporary world. In a society increasingly guided by scientific understanding and empirical knowledge, the integration of ethical thought with a naturalistic worldview becomes ever more relevant. As we embark on this exploration, we aim to unravel the complexities of naturalistic realism, critically examining its propositions, counterarguments, and practical implications in the vast tapestry of metaethics.

In the following sections, we will traverse the historical evolution of meta- ethics, define and dissect naturalistic realism, address its challenges, critique its propositions, and finally, illuminate its practical significance in our modern ethical landscape.

\section{Context and Evolution of Metaethics}

Before delving into the ancient beginnings of metaethics and its evolution, it's essential to understand the broader context and the significance of this philosophical journey. The history of metaethics is not just a chronicle of evolving ideas but a reflection of humanity's quest to understand the very basis of morality and ethical reasoning.

\subsection{The Significance of Metaethics in Human Thought}
Metaethics stands at the crossroads of philosophy, where inquiries about the nature of reality, human cognition, and moral principles converge. It goes beyond the mere assessment of ethical norms and delves into the deeper questions of what constitutes morality itself. This branch of philosophy grapples with questions about the origins of moral principles, the nature of moral judgments, and the meaning of ethical language. By exploring these fundamental aspects, metaethics seeks to provide a scaffold upon which all other ethical reasoning and moral philosophy can be structured.

The evolution of metaethical thought reflects the broader intellectual and cultural shifts in human history. Each era's approach to metaethics mirrors its prevailing attitudes towards knowledge, reality, and the place of humans in the universe. From the mystical and mythological interpretations of morality in ancient civilizations to the rational and empirical analyses of the modern age, the progression of metaethical thought maps onto the evolution of human understanding itself.

\subsection{The Early Stirrings of Ethical Inquiry}
The foundations of metaethical inquiry trace back to even before the well-documented philosophical discourses of Plato and Aristotle. Early civilizations grappled with questions of morality, often intertwined with mythology, religion, and cosmology. Ancient cultures, from the Egyptians to the Mesopotamians \footcite{Frankfort1948} and the Indus Valley \footcite[86-92]{Dales1962}, all had their conceptions of morality, often governed by divine laws and cosmic order.

In ancient Greece, the pre-Socratic philosophers, like Heraclitus and Py- thagoras, initiated the first forays into understanding the nature of reality, which indirectly laid the groundwork for ethical reasoning. Their explorations, though not explicitly ethical, broached fundamental questions about the universe, paving the way for more focused inquiries into human conduct and morality. For instance, Heraclitus of Ephesus (535–475 BC) was famous for his insistence on ever-present change as the fundamental essence of the universe, as stated in the famous saying “panta rhei”—everything flows \footcite[12-16]{Hessel2018}. Pythagoras of Samos (570–495 BC), known for his mathematical and scientific discoveries, also influenced philosophical thought, extending to ethics and the understanding of reality \footcite[111-121]{Lambek1981}.


\subsection{Ancient Foundations: Plato and Aristotle}
With this broader understanding of the significance and preludes to metaethical thought, we can now delve into the ancient beginnings of metaethical exploration. Also in ancient Greece, the pre-Socratic philosophers, like Heraclitus and Pythagoras, initiated the first forays into understanding the nature of reality, which indirectly laid the groundwork for ethical reasoning. Their explorations, though not explicitly ethical, broached fundamental questions about the universe, paving the way for more focused inquiries into human conduct and morality. With this broader understanding of the significance and preludes to metaethical thought, we can now delve into the ancient beginnings of metaethical exploration. The philosophical giants of ancient Greece, Plato and Aristotle, would take these early inquiries and shape them into more structured and enduring forms of ethical investigation, setting the course for the rich and complex journey of naturalistic realism in metaethics. They are the figures who laid the foundational stones of Western philosophical thought. 

Plato’s musings, particularly in dialogues like "Euthyphro," brought to light the dilemma of whether moral truths are divine commands or hold intrinsic value independent of such decrees \footcite[241-256]{Wiles2014}. His Theory of Forms postulated a realm of ideal, abstract entities, separate from the material world, where ethical ideals existed in perfection. This dualistic approach set the stage for centuries of philosophical inquiry into the nature of moral truths. 

Aristotle, diverging from his mentor, embarked on a path that would later resonate with the principles of naturalistic realism. In his "Nicomachean Ethics," Aristotle argued for a naturalistic foundation of virtue, rooting it in human nature and the pursuit of eudaimonia (flourishing or well-being) \footcite[121-147]{Leibowitz2013}. His teleological approach viewed ethics as deeply intertwined with the purpose and function of human life, a viewpoint that significantly influenced later naturalistic interpretations in metaethics.

\subsection{The Medieval to Early Modern Transition}
The medieval period saw the dominance of religious and theological perspectives in ethical thought, with figures like Thomas Aquinas integrating Aristotelian philosophy with Christian theology. Aquinas' moral philosophy, often referred to as Natural Law Ethics, is a synthesis of Christian theology, Aristotelianism, and Neoplatonism \footcite[126-142]{Stančienė2014}. He reinterpreted Aristotle’s ideas in a Christian context, emphasizing the importance of grace and divine law in moral reasoning \footcite[128-150]{Shanley2002}. 

However, the transition to the early modern era marked a pivotal shift. Philosophers began to increasingly question religious and supernatural explanations, favoring a more human-centered and rational approach to understanding morality. This period also saw the emergence of ethical theories like deontology and utilitarianism, which would later influence the development of naturalistic realism in metaethics.

\subsection{The Rise of Empiricism and Naturalism}
The Enlightenment era witnessed the rise of empiricism, emphasizing knowledge derived from sensory experience. This period saw a gradual shift towards a more naturalistic view of ethics. Philosophers like David Hume and later Jeremy Bentham and John Stuart Mill advanced the idea that ethical truths could be understood through human nature and empirical observation, laying the groundwork for modern naturalistic realism \footcite[298-350]{Skorupski2021}. Hume, in particular, challenged the notion that moral judgments could be derived from reason alone, emphasizing the role of human sentiments and the observable consequences of actions \footcite[1-11]{Ducheyne2017}.

\subsection{Twentieth Century: From Non-Naturalism to Naturalistic Realism}
The early twentieth century brought a resurgence of non-naturalist thought, particularly with G.E. Moore’s 'Principia Ethica,' which introduced the 'naturalistic fallacy' argument. Moore contended that ethical properties were indefinable and non-reducible to natural properties, a stance that significantly influenced metaethical discussions \footcite[291-311]{Baldwin2006}. 

However, the latter half of the century witnessed a revival of naturalistic realism. Philosophers like John Rawls and Philippa Foot rekindled interest in a naturalistic approach, advocating for a view of ethics that was grounded in human nature and social conditions. This period saw a growing consensus that ethical truths, while normative, were not disconnected from the empirical world.

\subsection{Conclusion of Section}
The historical journey of metaethics, from Plato's ideal forms to the naturalistic realism of the modern era, reflects a dynamic and evolving field. This rich tapestry of thought sets the stage for a deeper understanding of naturalistic realism's role in contemporary metaethics, highlighting its attempt to bridge the gap between ethical intuitions and the observable natural world.

Incorporating references from the provided research papers, here is the revised text with citations for the section "Defining Naturalistic Realism" in a discussion on metaethical theories:

\section{Defining Naturalistic Realism}

In the diverse landscape of metaethical theories, naturalistic realism stands out for its unique approach to understanding moral truths. This section aims to define and elucidate naturalistic realism, contrasting it with other metaethical perspectives, and highlighting its distinctive features \footcite[610-624]{Copp1991}.

\subsection{Essence of Naturalistic Realism}
At its core, naturalistic realism posits that moral facts and values are real and discernible within the natural world. It challenges the traditional divide between moral and empirical truths, advocating for a harmonious integration where moral properties are viewed as natural properties \footcite[532-547]{Dreier2018}. This perspective diverges significantly from ethical non-naturalism, which argues that moral truths exist independently of the natural realm, and from ethical subjectivism, which posits that moral truths are contingent on individual or cultural beliefs \footcite[843-851]{Joyce2014}.

Naturalistic realism is anchored in the belief that ethical statements can be true or false in the same way that factual statements about the physical world are. This stands in contrast to expressivism or emotivism, which view ethical statements as expressions of emotional attitudes or prescriptions for action rather than statements of fact \footcite[570-572]{FinlayCuneo2008}. 

\subsection{Key Tenets of Naturalistic Realism}
Empirical Foundation: Naturalistic realism maintains that ethical truths can be observed and understood through empirical investigation \footcite[339-362]{Stringer2018}. This approach underpins the idea that moral properties, such as 'goodness' or 'rightness', have naturalistic underpinnings that can be studied and understood in the context of human behavior, psychology, and social dynamics.

Moral Facts as Natural Facts: The theory asserts that moral facts are part of the natural order of things \footcite[154-176]{Brink2001}. It implies that moral judgments are not merely subjective or relative but are grounded in objective realities of the natural world.

Cognitivism: Naturalistic realism is a form of moral cognitivism, meaning it treats moral statements as capable of being true or false \footcite[]{McPherson2012}. This approach views ethical discourse as a domain of knowledge and truth, akin to scientific discourse.

\subsection{Differentiating from Other Theories}
While naturalistic realism shares certain features with other ethical theories, its unique blend of empiricism and moral objectivism sets it apart \footcite[471-490]{Arruda2018}. Unlike ethical relativism, which posits the variability of moral truths across different cultures, naturalistic realism argues for universal moral truths grounded in human nature and the natural world. Compared to deontological ethics, which focuses on duty and rules, naturalistic realism is more concerned with the natural properties that constitute moral truths.

\subsection{Conclusion of Section}
Naturalistic realism in metaethics presents a compelling framework for understanding ethics through the lens of naturalism and empiricism. It bridges the gap between moral philosophy and empirical science, offering a perspective where ethical truths are woven into the very fabric of the natural world. This theory not only invites philosophical contemplation but also encourages empirical exploration into the nature of morality.

\section{Challenges Addressed by Naturalistic Realism}

Naturalistic realism in metaethics does not just propose a framework for understanding moral truths; it also responds to several longstanding challenges in ethical theory. This section explores how naturalistic realism confronts these issues, offering insights into metaphysical, conceptual, epistemological, and semantic aspects of ethics.

\subsection{Metaphysical Challenge: The Place of Value in a World of Facts}
One of the primary challenges in metaethics is the metaphysical question of how ethical values coexist with a world largely understood in terms of empirical facts. Naturalistic realism addresses this by proposing that values are indeed part of the natural world. It rejects the dualism that separates facts and values, asserting that moral properties, such as goodness or rightness, are as real as physical properties like mass or charge. This view contends that moral values are not ethereal or supernatural but are grounded in the natural properties of beings, particularly humans. Naturalistic realism thus offers a unified vision of the world, where values are intertwined with the natural order.

\subsection{Conceptual Challenge: Relating Ought to Is}
The conceptual challenge in metaethics revolves around how normative statements (what ought to be) relate to descriptive statements (what is). Naturalistic realism strives to bridge this gap by showing that normative claims are deeply rooted in factual conditions. For instance, statements about what one ought to do are based on natural facts about human needs, social contexts, and the consequences of actions. This approach challenges the is-ought dichotomy popularized by David Hume, suggesting that ethical prescriptions are not disconnected from factual states but rather emerge from them.

\subsection{Epistemological Challenge: Justifying Ethical Beliefs}
Another significant challenge is epistemological: how can we justify our ethical beliefs, particularly when they often seem based on intuitions not easily verifiable through empirical means? Naturalistic realism approaches this by advocating for an empirical basis for ethical knowledge. It suggests that ethical beliefs can be justified through observation, scientific understanding of human nature, psychology, and sociology. This perspective posits that ethical intuitions are not arbitrary but are grounded in the natural experiences and evolutionary history of human beings.

\subsection{Connecting Thoughts and Words to Ethical Properties}
The semantic challenge in metaethics involves understanding how language and thought relate to ethical properties or facts. Naturalistic realism contends that ethical language successfully refers to real properties in the world. For instance, when we use terms like 'good' or 'just', we are referring to actual properties that can be observed and understood in the context of human interaction and societal norms. This theory holds that ethical discourse is not merely emotive or prescriptive but is descriptive of real, natural phenomena.

\subsection{Addressing Critiques and Counterarguments}
While naturalistic realism provides compelling responses to these challenges, it is not without its critiques. Critics argue that the theory oversimplifies the complexity of ethical phenomena, reducing rich, varied moral experiences to mere natural facts. They question whether naturalistic realism can adequately account for the normative force of ethical claims – how they guide, motivate, and bind us. In response, proponents of naturalistic realism argue that understanding moral phenomena in naturalistic terms does not diminish their richness or normativity. Instead, it grounds them in a reality that is shared and verifiable, providing a more robust foundation for ethical reasoning.

\subsection{Conclusion of Section}
Naturalistic realism in metaethics offers thoughtful and innovative solutions to some of the most pressing challenges in ethical theory. By advocating for a naturalistic basis for values, it provides a framework where ethical truths are not seen as separate from the empirical world but as an integral part of it. This perspective not only enriches our understanding of morality but also aligns ethical reasoning with our broader understanding of the natural world.

Incorporating references from the provided research papers, here is the revised text with citations for the section "Critique and Counterarguments" in a discussion on naturalistic realism:

\section{Critique and Counterarguments}

While naturalistic realism offers a compelling framework in metaethics, it is not without its criticisms. This section delves into the key critiques of naturalistic realism and the responses it garners, thereby providing a balanced view of this philosophical stance.

\subsection{Critique 1: The Naturalistic Fallacy Argument}
One of the most prominent critiques against naturalistic realism comes from G.E. Moore's concept of the "naturalistic fallacy." Moore argued that defining moral terms like 'good' in natural terms (such as pleasure or desire satisfaction) is fallacious. According to Moore, 'good' is a simple, indefinable quality, much like the color yellow; any attempt to define it in natural terms is misguided. This argument challenges the core of naturalistic realism, which seeks to ground ethical terms in natural properties \footcite[291-311]{Baldwin2006}.

Response: Proponents of naturalistic realism counter this by questioning Moore's assertion that moral properties are simple and indefinable. They argue that moral concepts are indeed complex and can be understood in terms of natural human experiences and societal interactions \footcite[532-547]{Dreier2018}.

\subsection{Critique 2: Over-Simplification of Ethical Phenomena}
Critics also argue that naturalistic realism oversimplifies the rich and varied nature of moral experiences by reducing them to natural facts. They claim that this reductionist view fails to capture the depth, complexity, and sometimes the paradoxical nature of ethical life \footcite[843-851]{Joyce2014}.

Response: Advocates of naturalistic realism argue that recognizing a natural basis for ethics does not necessarily lead to oversimplification. Instead, it provides a foundation for understanding the complexities of moral experiences within a framework that is coherent with our understanding of the natural world \footcite[610-624]{Copp1991}.

\subsection{Critique 3: Difficulty in Accounting for Normativity}
Another significant criticism is the challenge naturalistic realism faces in accounting for the normative aspect of ethics – the 'oughtness' or prescriptive power of moral claims. Critics question how natural facts can give rise to normative obligations and how moral imperatives can be derived from empirical observations \footcite[]{McPherson2012}.

Response: Naturalistic realists address this by arguing that normativity can arise from human nature and societal contexts. They propose that understanding the natural facts about human needs, social dynamics, and the consequences of actions can indeed inform what one ought to do. This approach does not deny the normative force of ethical claims but seeks to ground them in observable aspects of human life and interactions \footcite[339-362]{Stringer2018}.

\subsection{Conclusion of Section}
The critique of naturalistic realism highlights the ongoing philosophical debate and the complexity of grounding ethics in the natural world. While the theory faces challenges, particularly in explaining the normative force of ethical claims and avoiding the oversimplification of moral phenomena, its proponents offer robust responses. These responses strive to reconcile the empirical nature of moral facts with the rich, normative tapestry of ethical life.

\section{Practical Implications and Modern Examples}

The theoretical constructs of naturalistic realism in metaethics extend beyond philosophical discourse, influencing practical decision-making and moral reasoning. This section examines the real-world implications of naturalistic realism, highlighting its relevance in contemporary ethical dilemmas.

\subsection{Application in Ethical Decision-Making}
Naturalistic realism posits that ethical decisions can be informed by an understanding of natural human tendencies, societal norms, and empirical observations. This approach suggests that moral reasoning is not just an abstract exercise but is deeply rooted in practical realities. For example, in business ethics, naturalistic realism would advocate for decisions that align with both economic realities and ethical considerations derived from human nature and societal expectations. This could involve balancing profit-making with the natural human need for fairness and the social obligation to contribute positively to the community.

\subsection{Influence on Moral Education}
The naturalistic approach also has significant implications for moral education. By grounding moral principles in natural human experiences and social contexts, it offers a more relatable and tangible basis for teaching ethics. Educational programs informed by naturalistic realism would emphasize the development of moral reasoning based on empirical understanding of human behavior, social interactions, and the consequences of actions. This approach can foster a more practical and empathetic understanding of ethics among students.

\subsection{Contemporary Ethical Dilemmas}
Environmental Ethics: In the context of environmental challenges, naturalistic realism provides a framework for understanding our moral obligations towards the environment. It suggests that our ethical responsibility to protect the environment is not just a subjective preference but is grounded in the natural relationship between humans and the natural world. This perspective can guide policies and actions towards sustainable practices that are in harmony with the natural ecological balance.

Medical Ethics: The application of naturalistic realism in medical ethics can be seen in debates over issues like euthanasia and genetic engineering. By examining these issues through the lens of natural facts about human health, suffering, and the implications of medical interventions, naturalistic realism can offer grounded ethical guidelines that consider both the empirical realities of medical science and the normative aspects of patient well-being and autonomy.

Social Justice: Naturalistic realism also informs discussions on social justice, particularly in understanding the ethical basis for equality and fairness. It suggests that concepts like justice and equality are not just abstract ideals but are rooted in natural social dynamics and the inherent value of each individual. This perspective can drive policies and actions aimed at addressing inequalities and promoting a fairer society.

\subsection{Challenges in Practical Application}
Despite its strengths, applying naturalistic realism to real-world ethical dilemmas is not without challenges. One key issue is the complexity of translating theoretical principles into specific guidelines for action. Additionally, the diversity of human experiences and cultural contexts can make it difficult to derive universally applicable ethical norms solely from naturalistic premises.

\subsection{Conclusion of Section}
Naturalistic realism's emphasis on grounding ethics in the natural world offers a powerful tool for navigating contemporary ethical challenges. By connecting moral principles with empirical realities, it provides a framework that is both theoretically robust and practically relevant. While there are challenges in application, the naturalistic approach to ethics remains a valuable perspective for addressing the complex moral issues of our time.

\newpage
\section*{Conclusion}

\subsection*{Reflecting on the Journey Through Naturalistic Realism}
This essay has traversed the philosophical landscape of naturalistic realism in metaethics, exploring its foundational principles, addressing the challenges it faces, examining its critiques, and considering its practical implications. In concluding, we reflect on the key insights gained and ponder the future trajectory of this influential metaethical stance.

\paragraph{Summarization of Key Points}
Naturalistic realism offers a compelling perspective in metaethics by advocating that moral truths and values are integral to the natural world. It challenges the traditional separation of ethical values from empirical reality, proposing instead that moral facts are as observable and real as physical facts. This theory responds adeptly to the metaphysical, conceptual, epistemological, and semantic challenges in ethics, providing a framework where moral truths are grounded in natural human experiences and societal dynamics.

Despite its strengths, naturalistic realism faces critiques, particularly regarding the naturalistic fallacy and its ability to account for the normative force of ethical claims. However, its proponents provide robust counterarguments, emphasizing the complexity and diversity of moral experiences within a naturalistic framework.

The practical significance of naturalistic realism is evident in its application to contemporary ethical dilemmas. From environmental ethics to medical ethics and social justice, it offers a grounded approach to moral reasoning, balancing empirical understanding with normative considerations.

\paragraph{Final Thoughts on Naturalistic Realism in Metaethics}
As we look towards the future, naturalistic realism continues to hold substantial promise in enriching our understanding of ethics. Its alignment with the empirical and its insistence on the reality of moral facts make it particularly relevant in an age where scientific understanding profoundly influences our worldview. While challenges remain, the ongoing dialogue and refinement of naturalistic realism will undoubtedly contribute to a deeper and more nuanced understanding of the nature of ethics, its place in our lives, and its role in shaping a just and harmonious society.


\printbibliography

\end{document}
