\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, style=authortitle, citestyle=verbose-trad1]{biblatex}
\usepackage{biblatex}
\addbibresource{references.bib}

\title{Chapter 1: Roundabout Methods}
\author{Daniel Schellhorn}
\date{\today}

\begin{document}
\maketitle

\section{Main Theme/Topic of the Chapter}

The main theme of the chapter is the investigation of problem-solving behavior in animals, particularly focusing on their ability to find indirect or "roundabout" methods to reach a goal when the direct path is blocked.

The main theme of Chapter 1, titled "Roundabout Methods," from Wolfgang Köhler's book "The Mentality of Apes," is the problem-solving behavior of higher animals, particularly apes, when faced with obstacles that prevent them from taking a direct path to their goals. Köhler explores how these animals navigate around physical barriers to reach food or other objectives, which requires them to engage in indirect, more complex routes as opposed to straightforward, direct approaches.

\section{Wolfgang Köhler's Statements}

\subsection{Köhler's Observations and Conclusions}

\begin{itemize}
    \item Köhler observes that animals, including higher animals that rely on vision, tend to pursue goals directly unless obstacles prevent them. He suggests that the ability to find a roundabout route is not based on previous experience but rather on the animal's ability to perceive the entire field and understand the geometry of the situation to overcome obstacles.
    \item He notes that while this behavior is common and not particularly insightful in chimpanzees, it is not a given in lower animals. The ability to find a roundabout path is a significant cognitive achievement for them.
\end{itemize}

\subsection{Köhler's Statements on the Theme}

\begin{itemize}
    \item Köhler observes that higher animals, which rely on vision, typically approach their goals directly unless there are complications.
    \item He sets up experiments to force animals to take a roundabout route, thereby necessitating a more complicated movement geometry.
    \item Through these experiments, Köhler notes that chimpanzees, in particular, are capable of finding detours and can navigate complex paths to reach their objectives.
\end{itemize}

\subsection{Definitions and Terms}

\begin{itemize}
    \item Direct Way: The natural, biologically determined, straight-line approach an animal takes towards an objective.
    \item Roundabout Way: A more complex path that an animal must take when the direct way is blocked, requiring the animal to navigate around the obstacle to reach the objective.
    \item Obstruction: The physical barrier that prevents the direct approach.
    \item Geometry of Movement: The path or route an animal must conceptualize and follow to circumvent an obstacle and reach a goal.
    \item Surveyable situation: A scenario where the objective, obstruction, and possible detours are all visible to the animal.
\end{itemize}

\section{Current Research State and Relevant Papers}

\subsection{Current Research State}

Presently, research in animal cognition continues to explore the problem-solving abilities of various species, including apes. There is a focus on understanding the cognitive processes behind tool use, memory, spatial awareness, and the ability to learn from experience or observation. Researchers are also investigating the neural mechanisms that underlie these behaviors.

Studies often investigate the extent to which animals can understand cause and effect, plan for future events, and exhibit signs of consciousness or self-awareness.

\subsection{Brief Look at Current Relevant Papers}

\subsection*{Networked problem solving}
The paper "A theory of intelligence: networked problem solving in animal societies" \footcite{shour_2009} by Robert Shour explores the concept of collective intelligence in animal societies. The key points of the paper are as follows:

\begin{enumerate}
    \item \textbf{Collective Intelligence:} The paper discusses how the intelligence of a society emerges from the networking of individual intelligences and the accumulation of solved problems.
    
    \item \textbf{Economic Growth and Problem Solving:} Shour proposes a mathematical relationship between problem-solving and societal advancement, suggesting that economic growth is proportional to the network entropy of a society's population times the network entropy of the number of solved problems.
    
    \item \textbf{Network Entropy:} The concept of network entropy is central to Shour's theory, indicating the complexity and interconnectedness of societal problem solving processes.
    
    \item \textbf{Implications for Understanding Intelligence:} The paper provides a novel perspective on intelligence, emphasizing the role of social and collaborative aspects in the development of cognitive abilities in animal societies.
\end{enumerate}

\subsection*{Statistical Problems in Neurophysiology}

The paper "What to do if N is two?" \footcite{fries_maris_2021} by Pascal Fries and Eric Maris addresses the statistical methodologies used in in-vivo neurophysiology, particularly focusing on the implications of using small sample sizes in research studies. The key points of the paper are as follows:

\begin{enumerate}
    \item \textbf{Statistical Standards in Neurophysiology:} The traditional approach of pooling data from a small number of animals, often just two, is scrutinized for its effectiveness in providing reliable and generalizable conclusions.
    
    \item \textbf{Inference Limitations:} The authors argue that meaningful inferences about a population require larger sample sizes and a different statistical approach than what is commonly used. Studies with only a few animals limit the inference to those specific animals.
    
    \item \textbf{Ethical and Economic Considerations:} The paper discusses the challenges in conducting studies with larger numbers of animals, considering ethical and economic constraints. It suggests coordinated multi-center efforts for important questions.
    
    \item \textbf{Recommendations for Research Practices:} A reevaluation of statistical standards in neurophysiology research is recommended. For studies with small sample sizes, conclusions should be limited to the sample studied.
\end{enumerate} 

\newpage
\section{Teachings from the Chapter}

This chapter teaches us about the cognitive abilities of animals in problem-solving situations. It highlights the importance of understanding the mental processes that allow animals to perceive their environment, evaluate obstacles, and find alternative routes to reach their objectives. It also suggests that these abilities vary significantly across species and that what may be a simple task for one species can be a complex cognitive challenge for another. The chapter underscores the significance of experimental design in studying animal cognition and the need to differentiate between behaviors that result from genuine problem-solving versus those that occur by chance. Moreover, it implies that animals are not only driven by instinct but also possess the capacity for insight and learning, which allows them to navigate complex situations.

\end{document}
