\documentclass[12pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}

\begin{document}

\begin{letter}{Daniel Schellhorn \\ d.p.schellhorn@gmail.com \\ +49 177 2669996 \\ Sedanstr. 12, 49076, Osnabrück, Germany}

\opening{Dear Hiring Manager,}

With a strong academic background in mathematics and AI, and a deep passion for data science, I am excited about the opportunity to contribute to and learn from UBS’s innovative team.

The Data Science Internship at UBS offers a unique opportunity to engage with advanced analytics at the forefront of digital transformation in banking. I am particularly enthusiastic about creating abstract and strategic data models applicable across various domains. This abstraction process is fundamental in managing the complexity arising from integrating diverse types of data and structures. Specifically, the unifying data model of document databases like MongoDB, combined with enhanced Entity-Relationship (EER) diagrams, presents promising potential for comprehensive data conceptualizations.

I am currently completing dual Master’s degrees in Mathematics and Cognitive Science (with a focus on AI) at the University of Osnabrück, where I have developed a robust foundation in data science, statistics, and machine learning. My coursework and projects have equipped me with proficiency in Python, VS Code, GitHub, and statistical data science. Additionally, my Bachelor's degree in Mathematics from the University of Göttingen provided me with practical experience in numerical analysis and financial mathematics, aligning well with the skills required for this internship.

I am particularly drawn to UBS because of its commitment to driving digital transformation in banking and its focus on innovative analytics solutions. UBS’s dedication to remote work and embracing new work cultures, as highlighted in the discussion with Tom Savigar about the future of work, resonates with my interest in flexible and adaptive working environments. The shift towards work-habitats and digital nomadism aligns with my aspiration to work in dynamic and innovative settings.

Thank you for considering my application. I look forward to the opportunity to discuss how my background, skills, and enthusiasm align with the goals of your team. I am confident that my dedication and willingness to learn will make a positive impact at UBS.

\closing{Sincerely,}

\end{letter}

\end{document}
